---
title: "README"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "man/figures/README-"
)
```

# (SSMSE) Management Strategy Evaluation for Stock Synthesis (SS)

master:
[![Build Status](https://travis-ci.org/nmfs-fish-tools/SSMSE.svg?branch=master)](https://travis-ci.org/nmfs-fish-tools/SSMSE)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/nmfs-fish-tools/SSMSE?branch=master&svg=true)](https://ci.appveyor.com/project/nmfs-fish-tools/SSMSE)
[![codecov](https://codecov.io/gh/nmfs-fish-tools/SSMSE/branch/master/graph/badge.svg)](https://codecov.io/gh/nmfs-fish-tools/SSMSE)


**************

<https://nmfs-fish-tools.github.io/SSMSE/>


**************

## This is a repository for the Stock Assessment Tool: SSMSE
- Supported by the NOAA Fisheries Integrated Toolbox


## Disclaimer

“The United States Department of Commerce (DOC) GitHub project code is provided on an ‘as is’ basis and the user assumes responsibility for its use. DOC has relinquished control of the information and no longer has responsibility to protect the integrity, confidentiality, or availability of the information. Any claims against the Department of Commerce stemming from the use of its GitHub project will be governed by all applicable Federal law. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by the Department of Commerce. The Department of Commerce seal and logo, or the seal and logo of a DOC bureau, shall not be used in any manner to imply endorsement of any commercial product or activity by DOC or the United States Government.”

<!-- - This project code is made available through GitHub but is managed by NOAA at
 https://vlab.ncep.noaa.gov/redmine/projects/integrated-fisheries-toolbox/files -->

***** *******

## Installing the SSMSE R package

Note that the SSMSE is a work in progress and not yet a minimum viable product.

To install SSMSE from github:
```{r, eval = FALSE}
remotes::install_github("nmfs-fish-tools/SSMSE")
```
You can read the help files with

```{r, eval = FALSE}
?SSMSE
```

## An SSMSE toy example

Suppose we want to look at 2 scenarios, one where natural mortality (M) is specified correctly and one  where it is specified incorrectly in an estimation model (EM):
1. Cod model operating model (OM; M = 0.2) with cod model EM (assumed M = 0.2)
2. Lower M cod operating model (M = 0.1) with cod model EM (assumed M = 0.2)

Note that this is a toy example and not a true MSE, so there is not significantly different structure between the cod EM and OM, except for different M. We will assume we want to run the MSE loop for 6 years, with a stock assessment occuring every 3 years. The cod model's last year is 100, so the OM is initially conditioned through year 100. Then, after conditioning the operating model through year 100, assessments will occur in years 100, 103, and 106. Note that the assessment run in year 106 will generate future catch for years 107, 108, and 109, but these are not feed back into the operating model because the MSE loop is specified to only run through year 106).

First, we will load the `SSMSE` package and create a folder in which to run the example:
```{r, echo=FALSE, results="hide"}
devtools::load_all(".") # for rendering this readme when it is in the SSMSE pkg
```
```{r, eval=FALSE}
library(SSMSE) #load the package
library(r4ss) #install using remotes::install_github("r4ss/r4ss@development)
```
```{r}
# Create a folder for the output in the working directory.
run_SSMSE_dir <- file.path("run_SSMSE-ex")
dir.create(run_SSMSE_dir)
```

The cod model with M = 0.2 is included as external package data. However, we will need to modify it to use as an operating model with M = 0.1.
```{r}
cod_mod_path <- system.file("extdata", "models", "cod", package = "SSMSE")
# copy to a new location:
file.copy(from = cod_mod_path, to = run_SSMSE_dir, recursive = TRUE)
file.rename(from = file.path(run_SSMSE_dir, "cod"), to = file.path(run_SSMSE_dir, "cod-0.1"))
cod_0.1_path <- file.path(run_SSMSE_dir, "cod-0.1")
# make model read initial values from control file and not ss.par
start <- r4ss::SS_readstarter(file = file.path(cod_0.1_path, "starter.ss"), verbose = FALSE)
start$init_values_src # verify reading from the control file
# change the natural mortality paramter from 0.2 to 0.1 in the control files
r4ss::SS_changepars(dir = cod_0.1_path, ctlfile = "control.ss_new", 
              newctlfile = "control_modified.ss", strings = "NatM_p_1_Fem_GP_1", newvals = 0.1)
# remove files with M = 0.2
file.remove(file.path(cod_0.1_path, "control.ss_new"))
file.remove(file.path(cod_0.1_path, "control.ss"))
file.remove(file.path(cod_0.1_path, "ss.par")) # delete control file because no longer need.
# rename file with M = 0.1 to control.ss_new () and make a copy as the control file
file.rename(from = file.path(cod_0.1_path, "control_modified.ss"),
            to = file.path(cod_0.1_path, "control.ss"))
```

Rerun this model with no estimation to get valid ss.par and control.ss_new files:
```{r, results = "hide"}
# run SS with no estimateion
SSMSE:::run_ss_model(dir = cod_0.1_path, 
                     admb_options = "-maxfn 0 -phase 50 -nohess",
                     verbose = FALSE)
```

The argument `sample_struct` specifies the structure for sampling from the OM (and passing to the EM). The function `create_sample_struct` can be used to construct a simple sampling structure consistent with an input data file:

```{r}
EM_datfile <- system.file("extdata", "models", "cod", "ss3.dat", package = "SSMSE")
sample_struct <- create_sample_struct(dat = EM_datfile, nyrs = 6) # note warning
sample_struct
```
This sample_structure suggest that catch will be added to the estimation model every year (years 101 to 106), but an index of abundance (i.e., CPUE) and age composition (i.e., agecomp) will only be added in year 105. We can modify this sampling strategy however we would like.

Note that length comp (lencomp) includes an `NA` value for year. This is because
no consistent pattern was identified, so the user must define their own input.
In this case, we will remove sampling length comps all together:
```{r}
sample_struct$lencomp <- NULL # don't use length sampling
```

The same sampling structure will be used for both scenarios:
```{r}
sample_struct_list <- list("M-0.2" = sample_struct, "M-0.1" = sample_struct)

```
We can now use `run_SSMSE` to run the MSE analysis loop:

```{r, warning = FALSE, message = FALSE}
run_res_path <- file.path(run_SSMSE_dir, "results")
dir.create(run_res_path)
# run 1 iteration and 1 scenario of SSMSE using an EM.
run_SSMSE(scen_name_vec = c("M-0.2", "M-0.1"), # name of the scenario
          out_dir_scen_vec = run_res_path, # directory in which to run the scenario
          iter_list = list(1:2, 1:2), # run with 5 iterations each
          OM_name_vec = NULL, # specify directories instead
          OM_in_dir_vec = c(cod_mod_path, normalizePath(cod_0.1_path)), # OM files
          EM_name_vec = c("cod", "cod"), # cod is included in package data
          MS_vec = c("EM","EM"),       # The management strategy is specified in the EM
          use_SS_boot_vec = c(TRUE, TRUE), # use the SS bootstrap module for sampling
          nyrs_vec = c(6, 6),        # Years to project OM forward
          nyrs_assess_vec = c(3, 3), # Years between assessments
          rec_dev_pattern = c("none", "none"), # Don't use recruitment deviations
          impl_error_pattern = c("none", "none"), # Don't use implementation error
          sample_struct_list = sample_struct_list) # How to sample data for running the EM.
```

The function `SSMSE_summary_all` can be used to summarize the model results in a list of dataframes. Note that if you have issues, try reinstalling SSMSE using `remotes::install_github("nmfs-fish-tools/SSMSE")` and restarting your R session. Also, make sure you are using the development branch versions of [r4ss](https://github.com/r4ss/r4ss) and [ss3sim](https://github.com/ss3sim/ss3sim) (by installing `remotes::install_github("r4ss/r4ss@development")` and `remotes::install_github("ss3sim/ss3sim@development")`. These versions should be installed automatically when SSMSE is downloaded.

```{r}
# Summarize 1 iteration of output
summary <- SSMSE_summary_all(normalizePath(run_res_path))
```
Plotting and data manipulation can then be done with these summaries. For example, SSB over time by model can be plotted. The models include the Operating Model (cod_OM), Estimation model (EM) for the historical period of years 0-100 (cod_EM_init), the EM run with last year of data in year 103 (cod_EM_103), and the EM run with last year of data in 106 (cod_EM_106).
```{r plot_SSB}
library(ggplot2) # use install.packages("ggplot2") to install package if needed
library(tidyr) # use install.packages("tidyr") to install package if needed
library(dplyr)
summary$ts <- tidyr::separate(summary$ts,
                               col = model_run,
                               into = c(NA, "model_type"),
                               remove = FALSE,
                               sep = "_", 
                               extra = "drop")
# check values for cod_OM
summary$scalar %>%
  dplyr::filter(iteration == 1) %>% 
  dplyr::filter(scenario == "M-0.1") %>% 
  dplyr::select(iteration, scenario, NatM_p_1_Fem_GP_1, model_run)
  

# plot SSB by year and model run - need to correct using code from the 
# think tank
ggplot2::ggplot(data = subset(summary$ts, model_run %in% c("cod_OM", "cod-0.1_OM", "cod_EM_106")), 
                aes(x = year, y = SpawnBio)) +
                geom_vline(xintercept = 100, color = "gray") +
                geom_line(aes(linetype = as.character(iteration), color = model_type))+
                scale_color_manual(values = c("#D65F00", "black")) +
                scale_linetype_manual(values = rep("solid", 50))+
                guides(linetype = FALSE)+
                facet_wrap(~scenario)+
                theme_classic()
```

If you wish to delete the files created from this example, you can use:
```{r, eval = FALSE}
unlink(run_SSMSE_dir, recursive = TRUE)
```

## How can I contribute to SSMSE?

If you have thoughts about how to implement the [upcoming work](#roadmap-where-is-ssmse-headed-next) or are interested in helping develop SSMSE, please contact the developers by posting an issue in this repository or emailing nmfs.stock.synthesis@noaa.gov.

If you are interested in contributing, please read the [NMFS Fisheries Toolbox R Contribution Guide](https://github.com/nmfs-fish-tools/Resources/blob/master/CONTRIBUTING.md). This project and everyone participating in it is governed by the [NMFS Fisheries Toolbox Code of Conduct](https://github.com/nmfs-fish-tools/Resources/blob/master/CODE_OF_CONDUCT.md). By participating, you are expected to uphold this code. Please report unacceptable behavior to [fisheries.toolbox@noaa.gov](mailto:fisheries.toolbox@noaa.gov).

## Roadmap: Where is SSMSE headed next?

SSMSE is still a work in progress, with basic framework in development. Some new directions we hope to work on shortly:

 - Expanding on examples to illustrate the package
 - Improving usability of the wrapper functions that users access
 - Adding more complex sampling options
 - Adding functions to calculate performance metrics
 - Adding functions to make some basic plots of diagonstics and results

If you have thoughts about how to implement the upcoming work or are interested in helping develop SSMSE, please contact the developers by posting an issue in this repository or emailing <nmfs.stock.synthesis@noaa.gov>
