---
title: "README"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "man/figures/README-"
)
```

# (SSMSE) Management Strategy Evaluation for Stock Synthesis (SS)

master:
[![Build Status](https://travis-ci.org/nmfs-fish-tools/SSMSE.svg?branch=master)](https://travis-ci.org/nmfs-fish-tools/SSMSE)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/nmfs-fish-tools/SSMSE?branch=master&svg=true)](https://ci.appveyor.com/project/nmfs-fish-tools/SSMSE)
[![codecov](https://codecov.io/gh/nmfs-fish-tools/SSMSE/branch/master/graph/badge.svg)](https://codecov.io/gh/nmfs-fish-tools/SSMSE)


**************

<https://nmfs-fish-tools.github.io/SSMSE/>


**************

**This is a repository for the Stock Assessment Tool: SSMSE**

- Supported by the NOAA Fisheries Integrated Toolbox


**Disclaimer**

“The United States Department of Commerce (DOC) GitHub project code is provided on an ‘as is’ basis and the user assumes responsibility for its use. DOC has relinquished control of the information and no longer has responsibility to protect the integrity, confidentiality, or availability of the information. Any claims against the Department of Commerce stemming from the use of its GitHub project will be governed by all applicable Federal law. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by the Department of Commerce. The Department of Commerce seal and logo, or the seal and logo of a DOC bureau, shall not be used in any manner to imply endorsement of any commercial product or activity by DOC or the United States Government.”

<!-- - This project code is made available through GitHub but is managed by NOAA at
 https://vlab.ncep.noaa.gov/redmine/projects/integrated-fisheries-toolbox/files -->

***** *******

## Installing the SSMSE R package

Note that the SSMSE is a work in progress and not yet a minimum viable product.

To install SSMSE from github:
```{r, eval = FALSE}
remotes::install_github("nmfs-fish-tools/SSMSE")
```
You can read the help files with

```{r, eval = FALSE}
?SSMSE
```

## Troubleshooting Installation

Here are some tips:

- Make sure you are using the development branch versions of R packages `r4ss` and `ss3sim`. These can be installed separately using `remotes::install_github("r4ss/r4ss", ref = "development")` and `remotes::install_github("ss3sim/ss3sim", ref = "development")`.
- If R asks "Would you like to download from sources that need compilation?", select "no", as the older compiled versions should work fine.
- If R asks which packages you would like to update, select "none." Although it is good to update to keep packages current, install can sometimes be frustrating when updating many packages and skipping this step can eliminate issues in the short term.
- If running into an error during install, try closing out all R sessions open (e.g., other R GUI or R studio windows), restarting the R session and trying `remotes::install_github("nmfs-fish-tools/SSMSE")` again.
- If still running into an error during install, and it seems to be a problem installing a particular package, try restarting the R session, using `install.packages()` to download the package that caused the error, restarting the R session, and trying `remotes::install_github("nmfs-fish-tools/SSMSE")` again. This step may need to be done several times for different R packages.


Still having trouble installing SSMSE? Please don't hesitate to open an [issue](https://github.com/nmfs-fish-tools/SSMSE/issues).

## An SSMSE example

Suppose we want to look at how well we are able to achieve a specified management procedure under uncertainty in the operating model. We will look 2 scenarios, one where Steepness (H) is specified correctly and one  where it is specified incorrectly in an estimation model (EM):

1. **H-ctl**: Cod operating model (H = 0.65) with correctly specified cod model EM (fixed H = 0.65)
2. **H-1**: Cod operating model (OM; H = 1) with misspecified cod model EM (fixed H = 0.65)

Note that this is a simple example and not a true MSE, so the OM and EM structures for both scenarios are identical, except for different steepness between the OM and EM in scenario 2. We will assume we want to run the MSE loop for 6 years, with a stock assessment occuring every 3 years (and forecasting catch to maintain 40% of unfished spawning stock biomass). The cod model's last year is 100, so the OM is initially conditioned through year 100. Then, after conditioning the operating model through year 100, assessments will occur in years 100, 103, and 106. Note that the assessment run in year 106 will generate future catch for years 107, 108, and 109, but the future catch values are not input into the operating model because the MSE loop is specified to only run through year 106).

### Setup R workspace folders
First, we will load the `SSMSE` package and create a folder in which to run the example:
```{r, echo=FALSE, results="hide"}
devtools::install_local(".", upgrade = "never", quiet = TRUE, force = TRUE)
```
```{r, eval=TRUE}
library(SSMSE) #load the package
library(r4ss) #install using remotes::install_github("r4ss/r4ss@development)
library(foreach) #if using run_parallel = TRUE
library(doParallel) #if using run_parallel = TRUE
```

```{r}
# Create a folder for the output in the working directory.
run_SSMSE_dir <- file.path("run_SSMSE-ex")
dir.create(run_SSMSE_dir)
```
### Create the operating models (OMs)

The cod model with H = 0.65 (as in scenario 1) is included as external package data in SSMSE. However, we will need to modify it to use as an operating model with H = 1 (as in scenario 2). Note in this case that refit_OM is false, so the model is not being refit, just run through without fitting. To condition the new model on the same data as the input model, refit_OM should be TRUE.

```{r}
cod_mod_path <- system.file("extdata", "models", "cod", package = "SSMSE")

develop_OMs(OM_name = "cod", out_dir = run_SSMSE_dir, par_name = "SR_BH_steep",
            par_vals = 1, refit_OMs = FALSE, hess = FALSE)
# OM model for scenario 2
cod_1_path <- file.path(run_SSMSE_dir, "cod_SR_BH_steep_1")
```

### Examine the management procedure used

We will use the same management procedure for both scenarios:

1. Conduct a stock assessment every 3 years to get stock status.
2. Forecast from this stock assessment using the SS forecast file to get future catch.
3. Put this forecasted catch (without implementation error, in the case of this example) back into the OM. Extend the OM forward in time to get the true values for the population.

We will now take a look at step 2 in the management procedure, which is implemented using the forecasting module in SS. We will examine the forecast file for the estimation model to better understand how catches will be forecasted from the assessment. We will use the same management procedure in both of these scenarios, although for a full MSE analysis, it is likely that multiple management procedures would be compared.

```{r}
fore <- r4ss::SS_readforecast(
  system.file("extdata", "models", "cod", "forecast.ss", package = "SSMSE"),
  verbose = FALSE)
fore$Forecast 
fore$Btarget
```
`fore$Forecast = 3` means our forecasts from the assessment will use the F to achieve a relative (to unfished) spawning stock biomass. Based on `fore$Btarget`, the relative biiomass target is 40% of unfished spawning stock biomass. Note also that the control rule `fore$BforconstantF` and `fore$BfornoF` values are set low to make it unlikely that they will be used:

```{r}
fore$BforconstantF
fore$BfornoF
```
Futhermore, `fore$Flimitfraction` is set to 1 so that the forecasted catch is set equal to the overfishing limit (for simplicity):
```{r}
fore$Flimitfraction
```
Note that the number of forecast years is 1:
```{r}
fore$Nforecastyrs
```
However, an assessment will be conducted every 3 years and 3 years of forecasting is required. SSMSE will modify this value to the appropriate number of forecasting years.

More information on using the forecast module in SS to forecast catches is available in the [Stock Synthesis users manual](https://vlab.ncep.noaa.gov/web/stock-synthesis/document-library/-/document_library/0LmuycloZeIt/view/4513132).


### Adding observation error: Specify how to sample data from the Operating model
The argument `sample_struct` specifies the structure for sampling from the OM (and passing to the EM). The function `create_sample_struct` can be used to construct a simple sampling structure consistent with an input data file:

```{r}
datfile <- system.file("extdata", "models", "cod", "ss3.dat", package = "SSMSE")
sample_struct <- create_sample_struct(dat = datfile, nyrs = 6) # note warning
sample_struct
```
The sample structure specifies that catch will be added to the estimation model every year (years 101 to 106), but an index of abundance (i.e., CPUE) and age composition (i.e., agecomp) will only be added in year 105. We will use the same sampling scheme for both scenarios, but it is possible to specify different sampling for each scenario. The user could modify this sampling strategy (for example, maybe age composition should also be sampled from FltSvy 2 in Yr 102; the user could add another line to the dataframe in `sample_struct$agecomp`).

Note that length comp (lencomp) includes an `NA` value for year. This is because
no consistent pattern was identified, so the user must define their own input.
In this case, we will remove sampling length comps all together:
```{r}
sample_struct$lencomp <- NULL # don't use length sampling
```

The same sampling structure will be used for both scenarios:
```{r}
sample_struct_list <- list("H-ctl" = sample_struct, "H-1" = sample_struct)

```
### Adding process error through recruitment devitations

Process error can be added through the recruitment deviations. In this case, `rec_dev_pattern = "rand"` in the call to `run_SSMSE` is used to use random recruitment deviations with the same standard deviation as the historical recruitment deviation pattern. Set `scope = 2` so that the same recruitment deviation patterns are used across scenarios, but different patterns are use across iterations in the same scenario.

### Performance metrics

Quantitative performance metrics should be specified before conducting an MSE. Typically, a suite of performance metrics will be examined; however, for simplicity in this example, we will only look at what the achieved relative biomass was for the last 3 years of projection in the MSE to determine how it compares to the intended management target of 40% of unfished Spawning Stock Biomass. Note that we are only running our MSE projectsion for 6 years, but longer projections are typical in MSE analyses.

### Run SSMSE

Now, use `run_SSMSE` to run the MSE analysis loop (note this will take some time to run, ~ 20 min):

```{r, warning = FALSE, message = FALSE}
run_res_path <- file.path(run_SSMSE_dir, "results")
dir.create(run_res_path)
run_SSMSE(scen_name_vec = c("H-ctl", "H-1"),# name of the scenario
          out_dir_scen_vec = run_res_path, # directory in which to run the scenario
          iter_vec = c(5,5), # run with 5 iterations each
          OM_name_vec = NULL, # specify directories instead
          OM_in_dir_vec = c(cod_mod_path, normalizePath(cod_1_path)), # OM files
          EM_name_vec = c("cod", "cod"), # cod is included in package data
          MS_vec = c("EM","EM"),       # The management strategy is specified in the EM
          use_SS_boot_vec = c(TRUE, TRUE), # use the SS bootstrap module for sampling
          nyrs_vec = c(6, 6),        # Years to project OM forward
          nyrs_assess_vec = c(3, 3), # Years between assessments
          rec_dev_pattern = "rand", # Don't use recruitment deviations
          scope = "2", # to use the same recruitment devs across scenarios.
          impl_error_pattern = "none", # Don't use implementation error
          run_EM_last_yr = FALSE, # Run the EM in 106
          run_parallel = TRUE,
          sample_struct_list = sample_struct_list, # How to sample data for running the EM.
          seed = 12345) #Set a fixed integer seed that allows replication 
```
Here, only 5 iterations are run per scenario; However, in a real MSE analysis, running 100+ iterations would be preferred, so that the full range of uncertainty (given observation and process errors) is visible in the results.

### Summarize results

The function `SSMSE_summary_all` can be used to summarize the model results in a list of dataframes.

```{r}
# Summarize 1 iteration of output
summary <- SSMSE_summary_all(run_res_path)
```
Plotting and data manipulation can then be done with these summaries. For example, SSB over time by model can be plotted. The models include the Operating Model (cod_OM), Estimation model (EM) for the historical period of years 0-100 (cod_EM_init), the EM run with last year of data in year 103 (cod_EM_103), and the EM run with last year of data in 106 (cod_EM_106).

```{r results="hide"}
library(ggplot2) # use install.packages("ggplot2") to install package if needed
library(tidyr) # use install.packages("tidyr") to install package if needed
library(dplyr)
```
```{r plot_SSB}
# check values for cod_OM
summary$scalar %>%
  dplyr::filter(iteration == 1) %>% 
  dplyr::filter(scenario == "H-1") %>% 
  dplyr::select(iteration, scenario, SR_BH_steep, model_run)
  

# plot SSB by year and model run
ggplot2::ggplot(data = subset(summary$ts, model_run %in% c("cod_OM", "cod_SR_BH_steep_1_OM", "cod_EM_103")), 
                ggplot2::aes(x = year, y = SpawnBio)) +
                ggplot2::geom_vline(xintercept = 100, color = "gray") +
                ggplot2::geom_line(ggplot2::aes(linetype = as.character(iteration), color = model_run))+
                ggplot2::scale_color_manual(values = c("#D65F00", "black", "blue")) +
                ggplot2::scale_linetype_manual(values = rep("solid", 50)) +
                ggplot2::guides(linetype = FALSE) +
                ggplot2::facet_wrap(. ~ scenario) +
                ggplot2::theme_classic()
```

Now, calculate and plot the performance metric.
```{r plot_rel_SSB}
get_rel_SSB_avg <- function(summary, min_yr, max_yr) {
  OM_vals <- unique(summary$ts$model_run)
  OM_vals <- grep("_OM$", OM_vals, value = TRUE)
  B_unfished <- summary$scalar %>% 
                 filter(model_run %in% OM_vals) %>% 
                 select(iteration, scenario,SSB_Unfished)
  SSB_yr <- summary$ts %>% 
              filter(year >= min_yr & year <= max_yr) %>% 
              select(iteration, scenario, year, SpawnBio)
  SSB_yr <- left_join(SSB_yr, B_unfished) %>% 
              mutate(Rel_SSB = SpawnBio/SSB_Unfished) %>% 
              group_by(iteration, scenario) %>% 
              summarize(avg_SSB = mean(Rel_SSB), .groups = "keep") %>% 
              ungroup()
  SSB_yr
}
rel_SSB <- get_rel_SSB_avg(summary, min_yr = 104, max_yr = 106)

# function to summarize data in plot
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
ggplot(data = rel_SSB, aes(x = scenario, y = avg_SSB)) +
  geom_hline(yintercept = 0.4, color = "gray") +
  stat_summary(fun.data = data_summary, 
               position = position_dodge(width = 0.9), color = "blue") +
  scale_y_continuous(limits = c(0, 0.8)) +
  labs(title = "Long-term average relative SSB\n(years 101-106)", 
       x = "Scenario", y = "SSB/SSB_unfished") +
  theme_classic()
```

If you wish to delete the files created from this example, you can use:
```{r, eval = FALSE}
unlink(run_SSMSE_dir, recursive = TRUE)
```

## How can I contribute to SSMSE?

If you have thoughts about how to implement the [upcoming work](#roadmap-where-is-ssmse-headed-next) or are interested in helping develop SSMSE, please contact the developers by posting an issue in this repository or emailing nmfs.stock.synthesis@noaa.gov.

If you are interested in contributing, please read the [NMFS Fisheries Toolbox R Contribution Guide](https://github.com/nmfs-fish-tools/Resources/blob/master/CONTRIBUTING.md). This project and everyone participating in it is governed by the [NMFS Fisheries Toolbox Code of Conduct](https://github.com/nmfs-fish-tools/Resources/blob/master/CODE_OF_CONDUCT.md). By participating, you are expected to uphold this code. Please report unacceptable behavior to [fisheries.toolbox@noaa.gov](mailto:fisheries.toolbox@noaa.gov).

## Roadmap: Where is SSMSE headed next?

SSMSE is still a work in progress, with basic framework in development. Some new directions we hope to work on shortly:

 - Expanding on examples to illustrate the package (e.g., illustrating how performance metrics can be calculated)
 - Adding more complex sampling options
 - Adding functions to calculate performance metrics
 - Adding functions to make some basic plots of diagonstics and results

If you have thoughts about how to implement the upcoming work or are interested in helping develop SSMSE, please contact the developers by posting an issue in this repository or emailing <nmfs.stock.synthesis@noaa.gov>
